Cleaning Tropical Andes plant GBIF occurrence records

### 2023-08-01 code/L1/L1_TropicalAndes_plants_GBIF.Rmd
update overview and data output update flag record code add year cleaning add code to export data to csv

---
title: "Cleaning Tropical Andes plant GBIF occurrence records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: "TropicalAndes_GBIF_plant_occ.csv"
data output: "TropicalAndes_GBIF_plant_occ_cleaned.csv"
date: "2023-07-25"
output: html_document
---

@@

## Select columns of interest
```{r}
# Get column names
names(TropicalAndes_GBIF_plant_occ)
```

```{r}
# Subset to columns of interest
TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
  dplyr::select()
# remove records without coordinates
TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
  filter(!is.na(decimalLongitude)) %>%
  filter(!is.na(decimalLatitude))
```

```{r}
 #to avoid specifying it in each function
names(dat)[2:3] <- c("decimallongitude", "decimallatitude")

TropicalAndes_GBIF_plant_occ_clean <- TropicalAndes_GBIF_plant_occ%>%
  cc_val()%>%
  cc_equ()%>%
  cc_cap()%>%
  cc_cen()%>%
  cc_coun(iso3 = "countryCode")%>%
  cc_gbif()%>%
  cc_inst()%>%
  cc_sea()%>%
  cc_zero()%>%
  cc_outl()%>%
  cc_dupl()
```

>>>>

---
title: "Cleaning Tropical Andes plant GBIF occurrence records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Uses CoordinateCleaner to flag and remove problematic records from GBIF plant occurrence records."
data input: "TropicalAndes_GBIF_plant_occ.csv"
data output: "TropicalAndes_GBIF_plant_occ_cleaned.csv, TropicalAndes_GBIF_plant_occ_flagged.csv"
date: "2023-07-25"
output: html_document
---

@@

```{r}
# remove records without coordinates
TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
  filter(!is.na(decimalLongitude)) %>%
  filter(!is.na(decimalLatitude))
```

```{r}
flags <- clean_coordinates(x = TropicalAndes_GBIF_plant_occ,
                           lon = "decimalLongitude",
                           lat = "decimalLatitude",
                           countries = "countryCode",
                           species = "",
                           tests = c("capitals", "centroids", "countries",
                                     "duplicates", "equal", "gbif",
                                     "institutions", "seas","zeros"))
summary(flags)
plot(flags, lon = "decimalLongitude", lat = "decimalLatitude")
```

```{r}
#exclude problematic records
TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ[flags$summary,]
TropicalAndes_GBIF_plant_occ_flagged <- TropicalAndes_GBIF_plant_occ[!flags$summary,]
```

```{r}
table(TropicalAndes_GBIF_plant_occ_cleaned$year)
```

```{r}
# remove records from before second world war
TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ_cleaned %>%
  filter(year > 1945) 
```

# Write data to csv
```{r}
write.csv(TropicalAndes_GBIF_plant_occ_cleaned, file = file.path(output_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))

write.csv(TropicalAndes_GBIF_plant_occ_flagged, file = file.path(output_path,"TropicalAndes_GBIF_plant_occ_flagged.csv"))
```
